---
title: "Tesla Locations"
author: "Hernan Soulages"
date: "December 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The locations in the *locations.txt* file are from the [Tesla finder](https://www.tesla.com/findus#/bounds/49.38,-66.94,25.82,-124.39?search=supercharger,store,service,destination%20charger,&name=us). It was easy to extract the latitude and longitude but getting the information for type of location was tricky. For the purpose of this project, I just used a generic text for the popups. It would be nicer to have a list with the address and type of station and make the icons represent the type and the popups show the address.

```{r locations, echo=FALSE, warning=FALSE}
locations <- read.csv('locations.txt')
library(leaflet)
teslaIcon <- makeIcon(
#        iconUrl = 'https://www.tesla.com/sites/all/themes/custom/tesla_theme/assets/img/icons/favicon-16x16.png',
        iconUrl = 'https://www.tesla.com/sites/all/modules/custom/tesla_findus_map/assets/icon-supercharger.png?20150604',
        iconWidth = 16,
        iconHeight = 16
)
popupText <- 'Tesla location (store, service center, supercharger or destination charger)'
locations %>% leaflet() %>% addTiles() %>% 
        addMarkers(clusterOptions = markerClusterOptions(maxClusterRadius=30, disableClusteringAtZoom=10), icon = teslaIcon, popup = popupText)
```